A Survey of Exoskeleton Control Architectures and Algorithms\\ (Draft 3.0)

### 2015-09-29 exos/bleex.tex
more edits

keywords: sensitivity amplification; strength augmentation.\\

@@

The BLEEX sensitivity amplification control strategy models the torque applied by the human pilot on the exoskeleton as $d$ (assuming no outside disturbances).   Neglecting gravity, the exokeleton angular velocity is modeled as 
\[v = G r + S d ,\] 
where $G$ is the transfer function from actuator inputs ($G$ is the exoskeleton's dynamics), $r$ is the actuator input, and $S$ is the sensitivity or transfer function from human torque to exokeleton angular velocity.

@@

The goal is to maximize sensitivity to $d$ \emph{without direct measurement.}  Sensitivity amplification accomplishes this by creating a feedback loop from a controller, $C$, acting only on exoskeleton variables.  A new sensitivity equation,
\[S_{new} = \frac{v}{d} = \frac{S}{1 + G C} ,\]
is maximized by applying positive feedback.  To achieve a large sensitivity, BLEEX uses $C = (1-\alpha^{-1})G^{-1}$ so that $\alpha$ provides a direct (scalar) amplification factor. A low-pass filter is often added to the $C$ term in order to damp out high frequency dynamics of the exoskeleton, which are not captured in these models.  Note that the controller depends on an inverse dynamic model of the exoskeleton, $G^{-1}$.  Since the model is hybrid, these dynamics switch according to gait phases (single support, double support, stance).  BLEEX detects these transitions using foot sensors.  Assuming the single leg support dynamics are in the form,
\begin{equation}
\mM({\bf\theta})\ddot{{\bf \theta}} + \mC({\bf\theta},\dot{{\bf\theta}}) \dot{{\bf \theta}} + \mV({\bf\theta}) = {\bf T} + {\bf d} ,
\end{equation}
where ${\bf T}$ is a vector of actuator torques, the BLEEX control torque would be
\begin{equation}
{\bf T} = \mV({\bf\theta}) + (1 - \alpha^{-1}) \big [ \mM({\bf\theta})\ddot{{\bf \theta}} + \mC({\bf\theta},\dot{{\bf\theta}}) \dot{{\bf \theta}} \big ] .
\end{equation}
The user must provide the first torque component as the exoskeleton has no actuator acting between the foot and the ground (the system is underactuated).

@@

\end{refsection}

>>>>

% keywords: sensitivity amplification; strength augmentation.\\

@@

The BLEEX sensitivity amplification control strategy models the torque applied by the human pilot on the exoskeleton as ${\bf d}$ (assuming no outside disturbances).   Neglecting gravity, the exokeleton angular velocity is modeled as 
\[\vv = \mG \vr + \mS {\bf d} ,\] 
where $\mG$ is the transfer function from actuator inputs ($\mG$ is the exoskeleton's dynamics), $\vr$ is the actuator input, and $\mS$ is the sensitivity or transfer function from human torque to exokeleton angular velocity.

@@

The goal is to maximize sensitivity to ${\bf d}$ \emph{without direct measurement.}  Sensitivity amplification accomplishes this by creating a feedback loop from a controller, $\mC$, acting only on exoskeleton variables.  A new sensitivity equation,
\[\mS_{new} = \frac{\vv}{{\bf d}} = \frac{\mS}{1 + \mG \mC} ,\]
is maximized by applying positive feedback.  To achieve a large sensitivity, BLEEX uses $\mC = (1-\alpha^{-1})\mG^{-1}$ so that $\alpha$ provides a direct (scalar) amplification factor. A low-pass filter is often added to the $\mC$ term in order to damp out high frequency dynamics of the exoskeleton, which are not captured in these models.  Note that the controller depends on an inverse dynamic model of the exoskeleton, $\mG^{-1}$.  Since the model is hybrid, these dynamics switch according to gait phases (single support, double support, stance).  BLEEX detects these transitions using foot sensors.  Assuming the single leg support dynamics are in the form,
\begin{equation}
\mM({\bf\theta})\ddot{{\bf \theta}} + \mC({\bf\theta},\dot{{\bf\theta}}) \dot{{\bf \theta}} + \mV({\bf\theta}) = {\bf T} + {\bf d} ,
\end{equation}
where ${\bf T}$ is a vector of actuator torques and $\mM({\bf\theta})$, $\mC({\bf\theta},\dot{{\bf\theta}})$, and $\mV({\bf\theta})$ are the mass, Coriolis, and gravity terms (respectively), the BLEEX control torque would be
\begin{equation}
{\bf T} = \mV({\bf\theta}) + (1 - \alpha^{-1}) \big [ \mM({\bf\theta})\ddot{{\bf \theta}} + \mC({\bf\theta},\dot{{\bf\theta}}) \dot{{\bf \theta}} \big ] .
\end{equation}
The user must provide the first torque component as the exoskeleton has no actuator acting between the foot and the ground (the system is underactuated).

@@

\end{refsection}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../survey"
%%% End:

### 2015-09-29 exos/exosText_MIT.tex
more edits

\begin{figure}[thpb]
\centering
\includegraphics[width=4.in]{exos/figs/MIT/walkingCycle}
  \caption{}
  \vspace{-0.2in}
 \label{fig:walkingCycle} 
 \end{figure} 
The walking cycle around which the MIT control scheme was designed is shown in Figure \ref{fig:walkingCycle}.  The individual control strategies in the various phases of the walking cycle are highlighted in Figure \ref{fig:walkingCycle}.  For the hip there are four distinct strategies.  During the thrust phase, the hip exerts a torque that helps to raise the center of mass of the system.  During the extension spring phase, a virtual spring stiffness allows energy to be virtually stored while the center of mass moves forward.  During the swing assist phase the virtual energy is released, resulting in a torque being applied which assists in swinging the leg forward.  Finally, in leg retraction a torque is applied to help with foot placement and weight acceptance.  A diagram that illustrates how the sensors in the shank and hip joint are used to switch between these various phases is show in Figure \ref{fig:hipControl} 
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/hipControl}
  \caption{}
  \vspace{-0.2in}
 \label{fig:hipControl} 
 \end{figure}

 The knee controller in the MIT exoskeleton effectively functions independently from he hip controller.  During knee strike, the variable damper in the knee exerts a torque proportional to the rotational velocity of the knee joint.  A residual magnetic field remains in the knee joint after it is turned off, creating a resistive torque.  The knee therefore needs to be actively demagnetized at full extension during the last phase of stance to allow it to swing freely during the subsequent swing phase.  The state machine diagram that illustrates how the various phase transitions for the knee joint were implemented using the exoskeleton's on-board sensors is in Figure \ref{fig:kneeControl} 
  \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/kneeControl}
  \caption{}
  \vspace{-0.2in}
 \label{fig:kneeControl} 
 \end{figure}

@@

 \subsection{Opinion}

>>>>

\begin{figure}[thpb]
\centering
\includegraphics[width=4.in]{exos/figs/MIT/walkingCycle}
  \caption{}
  \vspace{-0.2in}
 \label{fig:walkingCycle} 
 \end{figure} 
The walking cycle around which the MIT control scheme was designed is shown in Figure \ref{fig:walkingCycle}.  The individual control strategies in the various phases of the walking cycle are highlighted in Figure \ref{fig:walkingCycle}.  For the hip there are four distinct strategies.  During the thrust phase, the hip exerts a torque that helps to raise the center of mass of the system.  During the extension spring phase, a virtual spring stiffness allows energy to be virtually stored while the center of mass moves forward.  During the swing assist phase the virtual energy is released, resulting in a torque being applied which assists in swinging the leg forward.  Finally, in leg retraction a torque is applied to help with foot placement and weight acceptance.  A diagram that illustrates how the sensors in the shank and hip joint are used to switch between these various phases is show in Figure \ref{fig:hipControl}.
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/hipControl}
  \caption{}
  \vspace{-0.2in}
 \label{fig:hipControl} 
 \end{figure}

 The knee controller in the MIT exoskeleton effectively functions independently from he hip controller.  During knee strike, the variable damper in the knee exerts a torque proportional to the rotational velocity of the knee joint.  A residual magnetic field remains in the knee joint after it is turned off, creating a resistive torque.  The knee therefore needs to be actively demagnetized at full extension during the last phase of stance to allow it to swing freely during the subsequent swing phase.  The state machine diagram that illustrates how the various phase transitions for the knee joint were implemented using the exoskeleton's on-board sensors is in Figure \ref{fig:kneeControl}.
  \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/MIT/kneeControl}
  \caption{}
  \vspace{-0.2in}
 \label{fig:kneeControl} 
 \end{figure}

@@

 \subsection{Assessment and Recommendations}

### 2015-09-29 exos/exosText_ihmc.tex
more edits

 Both IHMC MAE control approaches noted are relatively straight forward implementations of low-level torque control.  The first control scheme takes full advantage of the SEA actuators' ability to directly measure output torque and subsequently to perform closed-loop torque control, (see \ref{fig:IHMCTOR}). 
  \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/ihmc/torqueCon}
  \caption{}
  \vspace{-0.2in}
 \label{fig:IHMCTOR}   
 \end{figure}
 The controller takes in a desired torque, compares the desired value to the output torque measured at each individual actuator, then uses this value as well as its derivative to compute an actuator command signal, which in this case is a current.  The largest question in this controller, as presented in literature, is how to define the desired torque term $\tau_\text{des}$.

>>>>

 Both IHMC MAE control approaches noted are relatively straight forward implementations of low-level torque control.  The first control scheme takes full advantage of the SEA actuators' ability to directly measure output torque and subsequently to perform closed-loop torque control, (see Fig~\ref{fig:IHMCTOR}). 
  \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/ihmc/torqueCon}
  \caption{}
  \vspace{-0.2in}
 \label{fig:IHMCTOR}   
 \end{figure}
 The controller takes in a desired torque, compares the desired value to the output torque measured at each individual actuator, then uses this value as well as its derivative to compute an actuator command signal, which in this case is a current.  The largest question in this controller, as presented in literature, is how to define the desired torque term $\tau_\text{des}$.

### 2015-09-29 exos/exosText_roboKnee.tex
more edits

The RoboKnee device was built as a prototype system to enhance human strength, endurance and speed.  The main design objectoves in the RoboKnee system were to enhance human performance using a low-impedance mechanical system that incorporates a natural user interface, long life, and is comfortable to wear.  The main control objective of the suit is to provide a ``get out of the way" control scheme, wear the system inherently interacts with the user through a low-impedance interface.

The design of a linear series elastic actuator and a novel control method for controlling a single powered joint at the knee of the operator are presented in this section.

@@

The RoboKnee actuator was designed to specifically provide low impedance and high force feedback fidelity.  Series elastic actuators are a more robust option for measuring output forces and torques than using notoriously delicate and expensive load cells.

The actuator used in the RoboKnee device is shown in Figure \ref{fig:roboAct}.  The actuator consists of a drive train subassembly and an output carriage subassembly. The two subassemblies are coupled through die compression springs.  A servo motor directly drives a ball screw assembly to drive a ball nut flange which pushes on a spring retaining plate.  Secondary spring retaining plate is coupled to an output assembly.  Measuring the deflection in the die springs makes it possible to measure the force applied to the output of the linear SEA.  
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/roboKnee/roboSEA}
  \caption{}
  \vspace{-0.2in}
 \label{fig:roboAct}   
 \end{figure}

The linear actuator pictured in Figure \ref{fig:roboAct} { \bf weighs 1.13 kg, has a stroke length of 30.5 cm and diameter of 5.8 cm}.  The {\bf maximum speed of the actuator is 28 cm/s}.  The actuators {\bf maximum continuous force is 565N, with a peak maximum force of 1330 N}, with {\bf maximum continuous power of 164 W, and maximum peak power of 634 W}.

The force control bandwidth of the RoboKnee actuator is dependent on the magnitude of the force, as significant time delays can occur for larger forces due to the fact that there is an inherent time delay due to the time it takes for the spring to compress.  Thus, the {\bf small force control bandwidth of the system is 35 Hz} and the {\bf high force bandwidth is around 7.5 Hz}.

@@

 The RoboKnee mechanism consists of a single actuated degree of freedom at the flexion/extension joint of the operator.  Note that the mechanism itslef does not provide a pathway for transferring weight from a load directly to the ground.  The mechanism transfers loads directly to the users musculoskeletal structure, and focuses primarily on augmenting the torque supplied to the users knee joint.

 The RoboKnee mechanism itself is composed of an off-the-shelf knee brace modified with structural pieces to extend the brace and provide actuator attachment points.  The RoboKnee mechanism is shown in \ref{fig:roboKnee}.  
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/roboKnee/roboBrace}
  \caption{}
  \vspace{-0.2in}
 \label{fig:roboKnee}   
 \end{figure}
 There exact length between attachment points of the actuator is not listed in the literature, but it appears that the attachment points are at minimum the stroke displacement of the actuator  away from each other, so a {\bf minimum moment arm of 30.5 cm} is a reasonable approximation.

 The RoboKnee mechanism uses a linear potentiometer spanning the spring retaining plates to measure displacement in the springs and subsequently to measure force at the actuator output.  The forces between the users foot and the ground are also measured using single axis load cells.  Potentiometers which measure the knee joint angle are also employed in this design.

@@

 The control scheme for the RoboKnee device is a hierarchical approach, which uses a straightforward mid-level force generation scheme coupled with a low-level closed-loop force-based control loop.  The low-level, or rather joint level control is a PD control loop.

 The mid-level controller of the RoboKnee uses several simplifying assumptions to perform ``force amplification" to offload a percentage of the torque required by the user to actuate the system's knee joint.  This is achieved through a positive force feedback amplification which is based on measuring the ground reaction force and using this value to calculate the torque that would be required to produce this force in a static situation, namely, \[ \tau = R \times F,\] where $R$ is the vector from the ground reaction force to the users knee and $F$ is the ground reaction force.  Note that in the RoboKnee implementation this value can only be estimated due to the fact that the ankle joint angle and the ground reaction force is assumed to be purely vertical (which is a result of the single axis force sensors on the users feet).

 Once the estimated joint torque $\tau$ is calculated, an amplification factor is used in the RoboKnee control scheme to decide how much of the required knee torque will be provided by the control system.  For example, if the amplification factor were chosen to be one the actuator would completely compensate for the sensed ground reaction forces, and if it were set to zero to exoskeleton would provide no force.

 \subsection{Opinion}

@@

 The overall design of the RoboKnee mechanism is most likely not the greatest, as it significantly increases the effective volume of the operator-system leg, but as a prototype this concept offers interesting benefits.

>>>>

The RoboKnee is a prototype system designed to enhance human strength, endurance and speed.  To achieve these goals, RoboKnee uses a low-impedance mechanical system that incorporates a natural user interface, long life, and is comfortable to wear.  The suit's controller implements a ``get out of the way'' control scheme, where the system interacts with the user through a low-impedance interface.  This section focuses on RoboKnee's linear series elastic actuator design and the novel control method the device employs to coordinate its single powered knee joint.

@@

The RoboKnee actuator is specifically designed to provide low impedance and high force feedback fidelity.  The device uses a series elastic actuator to obtain robust output force measurements.  The series elastic actuation scheme allows RoboKnee to avoid load cells, which are both delicate and expensive.

The actuator used in the RoboKnee device (see Figure \ref{fig:roboAct}) consists of a drive train and output carriage. These two sub-assemblies are coupled through die compression springs.  A servo motor directly connected to a ball screw assembly drives a ball nut flange that pushes on a spring retaining plate.  A secondary spring retaining plate is coupled to an output assembly.  By measuring the deflection in the die springs, RoboKnee is able to obtain the force applied to the output of the linear SEA.  
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/roboKnee/roboSEA}
  \caption{}
  \vspace{-0.2in}
 \label{fig:roboAct}   
 \end{figure}

The linear actuator pictured in Figure \ref{fig:roboAct} { \bf weighs 1.13 kg, has a stroke length of 30.5 cm and diameter of 5.8 cm}.  The {\bf maximum speed of the actuator is 28 cm/s}.  The actuator's {\bf maximum continuous force is 565N and its maximum peak force is 1,330 N}, with a {\bf maximum continuous power of output of 164 W, and maximum peak power of 634 W}.

The force control bandwidth of the RoboKnee actuator is dependent on the magnitude of the force.  Larger forces can induce significant time delays due to the time it takes for the spring to compress.  Thus, the {\bf small force control bandwidth of the system is 35 Hz} and the {\bf high force bandwidth is around 7.5 Hz}.

@@

 The RoboKnee mechanism consists of a single actuated degree of freedom powering the flexion/extension of the operator's knee joint.  Note that the mechanism itself does not provide a pathway for transferring weight from a load directly to the ground.  The mechanism transfers loads directly to the users musculoskeletal structure, and focuses primarily on augmenting user knee joint torques.

 The RoboKnee device is composed of an off-the-shelf knee brace modified with structural pieces to extend the brace and provide actuator attachment points (see Figure~\ref{fig:roboKnee}).
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/roboKnee/roboBrace}
  \caption{}
  \vspace{-0.2in}
 \label{fig:roboKnee}   
 \end{figure}
 The exact length between actuator attachment points is not listed in the literature, but it appears as though the attachment points are at a minimum of one actuator stroke displacement away from each other. Thus, we assume a {\bf minimum moment arm of 30.5 cm} is a reasonable approximation.

 RoboKnee uses a \textbf{linear potentiometer} spanning the spring retaining plates to measure displacement in the springs and subsequently to measure force at the actuator output.  The forces between the users foot and the ground are also measured using \textbf{single axis load cells}.  Potentiometers which measure the knee joint angle are also employed in this design.

@@

 RoboKnee employs a hierarchical control strategy, which uses a straightforward mid-level force generation scheme coupled with a low-level closed-loop force-based control loop.  The low-level (joint level) control is a PD control loop.

 The RoboKnee's mid-level controller uses several simplifying assumptions to perform ``force amplification'' to offload a percentage of the torque required by the user to actuate the system's knee joint.  This is achieved through positive force feedback amplification.  The approach uses the measured ground reaction force to calculate the torque that would be required to produce this force in a static situation, \[ {\bf \tau} = \vR \times \vF.\]  The $\vR$ term represents the vector from the ground reaction force to the user's knee and $\vF$ is the ground reaction force.  Note that in RoboKnee's implementation, $\vF$ can only be estimated due to the fact that the ankle joint angle and the ground reaction force is assumed to be purely vertical (a limitation induced by the single axis force sensors on the user's feet).

 Once the estimated joint torque, ${\bf \tau}$, is calculated, an amplification factor specifies how much of the required knee torque will be provided by the control system.  For example, with unity amplification factor, the actuator would completely compensate for the sensed ground reaction forces.  Similarly, with zero amplification, the exoskeleton provides no force.

 \subsection{Assessment and Recommendations}

@@

 The overall design of the RoboKnee mechanism is most likely not the greatest, as it significantly increases the effective volume of the operator-system leg, but as a prototype, this concept offers interesting benefits.

### 2015-09-29 survey.tex
more edits

\newcommand{\vF}{{\bf F}}
\newcommand{\vV}{{\bf V}}
\newcommand{\vG}{{\bf G}}
\newcommand{\vT}{{\bf T}}

@@

I copied figure captions and numbers. Matt did not.  Coordinate?

@@

I don't point out what the terms are in the dynamics equations.

Discuss IHMC -- issues with high impedance torque control in practice.

Talk a lot about drawbacks of SEAs in IHMC.  Are these included in MIT?

>>>>

\newcommand{\vF}{{\bf F}}
\newcommand{\vV}{{\bf V}}
\newcommand{\vG}{{\bf G}}
\newcommand{\vT}{{\bf T}}
\newcommand{\vR}{{\bf R}}

@@

@@

Discussion blurb -- all these exos either measure user intention directly and have sensor noise and calibration challenges, or measure at the suit and perform amplification.  The amplification strategy is highly susceptible to outside disturbances and therefore not appropriate in combat situations.
Combine the two with filtering or directly address the sensing challenge.

Discussion blurb (pertains to IHMC and probably others) -- issues with high impedance torque control in practice.

Talk a lot about drawbacks of SEAs in IHMC.  Are these included in MIT / RoboKnee?  (low control bandwidth is mentioned in RoboKnee).  
We should have a blurb about the strengths/weakness of SEAs and possible options -- stiffer springs for force measurement and spring emulation for compliant control?
